### 2016-04-06 ms/ms_probaweb.tex
Quasi final version

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{color,lineno,setspace,multirow}
\usepackage[sc]{mathpazo}
\usepackage{graphicx}
\usepackage[top=2.4cm,left=2.4cm,top=2.4cm,bottom=2.4cm,includefoot]{geometry}
\usepackage{rotating}

@@

\textbf{Authors:} Dominique Gravel$^{1,2,*}$, Ben Baiser$^{3}$, Jennifer A. Dunne$^{4}$, N\'eo
Martinez$^{5}$, Tommy Nyman$^{6}$, Timoth\'ee Poisot$^{2,7}$, Tomas Roslin$^{6}$, Spencer Wood$^{8}$, Daniel B. Stouffer$^{9}$, Jason Tylianakis$^{9}$\\

@@

3: University of Canterbury at Christchurch, School of Biological Sciences,\\

@@

5: \\

@@

@@

@@

Community ecology is defined in most textbooks as \emph{the study of the
interactions that determine the distribution and abundance of organisms}
(Krebs2001). Despite a general concensus on this definition (Scheiner2007), it
is surprising that most research on the variation of community structure has
focused mostly on the turnover of species composition (Anderson2011),
neglecting variation in the way they interact with each other (McGill2006).
Not surprisingly given this fact, biogeographers are still struggling to
figure out if interactions do impact species distribution (Wisz2012;
Kissling2012). There has been a step ahead with recent methodological
improvements accounting for interactions in species distribution models
(Pollock2014; Pelissier2013), but those remain nonetheless a 'species- based'
approach to communities where interactions are fixed covariates affecting
distribution. Here we plea for a more integrated approach of species and
interaction distribution.

The problem of community assembly is often formulated as \emph{how do we
sample a regional pool of species to constitute a local community?} This
question could be rewritten to adress the problem of network assembly, as
\emph{how do we sample a regional pool of interactions to constitute a local
interaction network?}. An illustration of this problem for a food web is
provided at Figure 1. The metaweb represents potential interactions among all
species that could be found in a given area. In this particular case, there
are 274 nodes, and 1077 links among plants, herbivores and parasitoids from
Northern Europe. An instance of a local community is also illustrated, with 44
nodes and 160 interactions. Only $51.2\%$ of all potential interactions are
realized. Our objective here is to provide a conceptual framework to explain
the sampling of the regional pool of interactions, along with a quantitative
method to predict it. The problem could be formalized by sequentially
understanding why only a fraction of the species are locally co-occurring, and
after why these species are interacting or not.

There are multiple causes to spatial turnover in community composition. The
first and most studied driver is the effect of variation in the abiotic
environment on species performance. Combined with specific responses in
demography, it generates variation among localities by locally selecting the
fittest species (Leibold2004). Stochasticity additionnaly plays a role, either
because of the inherently unpredictable nature of colonization and extinction
events, or because of strong non-linear feedbacks generating alternative
transients and equilibrium (Chase2007; Vellend2014). The analysis of community
turnover is usually performed with data represented in a table with rows
corresponding to sites (or measurements) and the columns to species. Beta-
diversity metrics quantify the variance of this community data (Legendre2005).
Traditionnal approaches rely on measures of the dissimilarity between
communities, using indices such as the Jaccard or the Bray Curtis measures of
dissimilarity. Recent methods decompose the total variation of the community
data into species and site contributions to beta diversity (Legendre2013).
Even though these methods compare whole list of species between sites, or
measurements, they remain fundamentally 'species-based' since they are all
based on the within column variation. None of them consider the variation of
associations (pairs or higher order motifs) explicitly.

The niche is by far the dominant concept to explain species distribution and
community assembly, from the local to the global scale. Following Hutchinson,
the niche is viewed as the set of environmental conditions allowing a
population to establish and maintain a population (see also Holt2009).
Community turnover arises following successive species replacement along an
environmental gradient, in agreement with a Gleasonian view of communities
(Gleason). The concept is straighforward to operationalize with species
distribution models, as it maps naturally on the available data (both
distribution and environmental data) and a vast array of statistical tools
representing it (e.g. Biomod, MaxEnt). It is however much harder to account
for ecological interactions in this approach (Peterson2011). These are often
viewed as externalities, constraining or expanding the range of environmental
conditions required for a species to maintain a population (Pulliam2000;
Soberon2007).

The network approach proposes a convenient formalism to represent the
structure of local communities. Species are represented as nodes and
interactions by links. The data could also be represented by matrices, with
each species in rows and columns, and the entries representing the occurrence
or the intensity of an interaction. Studies of network diversity are mostly
concerned by the distribution of the interactions within a location and not so
much by the variation among locations (Dunne2005; Bascompte2007; Ings2007;
Kefi2012). Network complexity is computed as the number of interactions in the
case of binary networks or interaction diversity in the case of quantitative
networks (Bersier2002). However, there is now evidence that ecological
interactions do also vary in space and time (Poisot2012; Trojelsgaard2015).
The variability of community structure in this situation arises from the
turnover of species composition, along with the turnover of the interactions
among pairs of species. The occurrence and intensity of interactions could
vary because of the environment, species abundance and higher order
interactions (Poisot2015a). The variation in community composition is often
independent of the variation of ecological interactions, suggesting these two
components of network variability respond to different drivers (Poisot2012).

Interestingly, the ecological network literature also has its own 'niche model' to
position a species in a community (Williams2000). The niche of a species in
this context represents the multidimentional space that could represent all of
its interactions. Each species is characterized by a niche position, an
optimum and a range over 3 to 5 different niche axes (Williams2000;
Eklof2013). The niche model has been successful at explaining the complexity
of a variety of networks, from food webs to plant-pollinator systems. The
conceptual framework is however limited to local communities and does not
provide any explanation to the turnover of network structure along
environmental gradients.

Here we adopt the view that a community structure is best represented as an
ecological network of interactions and develop a theory to explain its
turnover in space and time. We propose a new description of the niche that
integrates the effect of the environment on species distribution and on
ecological interactions. We first present the conceptual framework and then
formalize it mathematically with a probabilistic approach to the sampling of
the regional pool of interactions. We apply the framework to study the spatial
variation of host-parasite interactions across Europe. We find that the
variation of the environment causes both species and interaction turnover. The
network structures changes systematically across the latitudinal gradient,
with a peak of connectance at intermediate latitudes. At the pairwise level,
the statistical approach could be conceived as an interaction distribution
model. At the community level, the approach provides a likelihood based
method to compare different hypotheses of network turnover.

@@

Correctly describing the niche is key to understand turnover in community
structure. Despite several attempts to refresh the conceptual basis of what
ecological niches are, ecologists have not moved far past the "n-dimensional
hypervolume" formalism introduced by Hutchinson. Despite its
intuitive interpretation and translation into species distribution models
(Boulangeat2012; Blonder2014), the concept has been constantly criticized
(Hardin1960; Peters1991; Chase2003; Silvertown2004; Soberon2007) and several
attempts have been made to expand and reinforce it.

Part of the problem surrounding the definition of the niche has been clarified
with the distinction between Eltonian and Grinnellian definitions (Chase2003).
The Grinnellian dimension of the niche is the set of environmental conditions
required for a species to maintain a population in a location. The Grinnellian
niche is the most intuitive to apply and is the conceptual backbone of species
distribution models. The Eltonian niche on the other hand is the effect of a
species on its environment. This aspect of the niche is well known by
community ecologists, but is trickier to turn into predictive models.
Nonetheless, the development of the niche model of food web structure
(Williams2000) and its parameterization (Williams2010; Gravel2013) made it
more operational.

These perspectives are rather orthogonal to each other and thus lead to
considerable confusion in the literature (McIntyre). Chase2003 attempted to
reconcile them in their definition of the niche: \textit{[The niche is] the
joint description of the environmental conditions that allow a species to
satisfy its minimum requirements so that the birth rate of a local population
is equal or greater than its death rate along with the set of per capita
effects of that species on these environmental conditions.} Their
representation merges zero-net growth isoclines figuring the Grinnelian niche
(when do the population persists) with impact vectors figuring the Eltonian
niche (what is the per capita impact). While this representation has been very
influential in community ecology at the local scale (the resource-ratio theory
of coexistence - Tilman1982), it remains impracticable at the large spatial
scale because of the difficulties to measure it. The absence of any
mathematical representation of the niche that could easily be fit to
ecological data perhaps explain why biogeographers are still struggling to
develop species distribution models taking into account ecological
interactions.

We propose to integrate the two perspectives of the niche with a visual
representation of both components. The underlying rationale is that, in
addition to the environmental constraints on demographic performance, any
organism requires resources to sustain its metabolic demand and reproduction.
Abiotic environmental axes are any non-consumable factors affecting the
demographic performance of an organism. Alternatively, the resource axes are
traits of the resources allowing interactions with the consumer. The niche
should therefore be viewed as the set of abiotic environmental factors (the
Grinnelian component) along with the set of traits (the Eltonian component)
allowing a sustainable population to establish and maintain at a location.
Accordingly, each species could be characterized by an optimal position in
both the environmental (x-axis) and the trait (y-axis) plane. The integrated
niche is then the hypervolume where interactions could occur and sustain a
population. This approach radically change the representation of the niche,
putting species distribution and ecological interactions in the same
formalism.

The limits of the niches could be independent of each other (as in the example
at Fig. 2), alternatively interact. For instance, the optimimal prey body size
for predatory fishes could reduce with increasing temperature (Lelong2015),
which would make diet boundaries functions of the environment. The other way
around, we could also consider that the growth rate of the predator could
change with the body size of the preys it feeds on, thereby altering the
environmental boundaries.

\newpage
%========================================================%
\section*{A probabilistic representation of ecological interactions networks in space}

We now formalize the integrated niche with a probabilistic approach to
interactions and distribution. We seek to represent the probability an
interaction between species $i$ and $j$ occurs at location $y$. We
define $L_{ijy}$ as a stochastic variable and are looking
at the probabilty this event occurs, $P(L_{ijy})$. The occurrence of an
interaction is dependent on the co-occurrence of species $i$ and $j$. This
argument might seem trivial at first, but the explicit consideration of this
condition in the probabilistic representation of ecological interactions will
prove fundamental to understand their variation. We thus define $X_{iy}$ as a
stochastic variable representing the occurrence of a species $i$ at location
$y$, and similarly $X_{ijy}$ the co-occurrence of species $i$ and $j$. The
quantity we seek to understand is the probability of a joint event:

@@

Or simply said, the probability of observing both species $i$ and $j$, and
an interaction from $i$ to $j$. This probability could be decomposed
in two parts using the product rule of probabilities:

@@

The left term is the probability of observing the two species co- occurring at
location $y$. It corresponds to the Grinnelian dimension of the niche. The
right term is a conditional probability, representing the probability that an
interaction occurs between species $i$ and $j$, given their set of traits
$\mathbf{T}$ and they are co- occurring. It is referred as the metaweb and
corresponds to the Eltonian dimension of the niche described above. We will
see below how this formalism could be directly fitted to empirical data. But
before turning to an application, we will discuss the interpretation of
different variants of these two terms.

@@

There are several variants to the co-occurrence probability representing
different hypotheses about the temporal and spatial variation in network
structure (see the explicit formulations at Table 1). The simplest model
relates co- occurrence probability directly to the environment, $P(X_{ijy} |
E_y)$. In this situation there is no underlying assumption about the
ecological processes responsible for co-occurrence. It could arise because of
the impact of ecological interactions on distribution (Pollock2014) or
alternatively because of share environmental requirements. In the former case,
species are not independent to each other and the conditional occurrence must
be accounted for explicitlty, $P(X_{ijy} | E_y) = P(X_i | E_y, X_j)P(X_j |
E_y)$. In the later case, species are independent and only the marginal
occurrence must be accounted for, $P(X_{ijy} | E_y) = P(X_i | E_y)P(X_j |
E_y)$.

The co-occurrence probability itself could be dependent on ecological
interactions. Direct pairwise interactions such as competition, facilitation
and predation have long been studied for their impact on co-distribution (such
as in the cases studied by Diamond1976, Connor1980, Gotelli2000 etc..). Second
and higher order interactions (e.g. trophic cascade) could also impact co-
occurrence. Co-occurrence in ecological networks is however a topic of its
own, influenced by the degree distribution and species richness
(Cazelles2015). Almost only first order and second order interactions do
impact co-occurrence. The covariance of interacting species to an
environmental gradient also influences co-occurrence (Cazelles2016). Because
of the complexity of relating co-occurrence to the interaction network
structure, we  will focus here on the variation of interactions and not on the
distribution, and leave this issue for Discussion and future research.

@@

There are also variants of the metaweb. First, most documented metawebs have
thus far considered that ecological interactions are deterministic, not
probabilistic (e.g. Havens1992; Woods2015). Species are assumed to interact if
they are found together in a location, independently of their abundance and
the environment. In other words, $P(L_{ijy}) = 1$ if $X_{ijy} = 1$, and 0
otherwise. This approach might be a reasonnable approximation when the
sampling and inference scales are large enough so that probabilities of
observing at least one interaction converges to unity and that the only
variation of networks considered arises from species distribution.

Ecological interactions could also vary with the environment, such as
$P(L_{ijy} | E_y)$. Although it is not common to see a conditional
representation of ecological interactions, experimental studies of pairwise
interactions revealing their sensitivity to the environment are common (REF).
For instance, it has been documented that the predation risks of shorebirds do
vary at the continental scale, from the south to the north (REF). The effect
of the environment on interactions propagate up the community and influence
network structure (REF).

>>>>

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{color,lineno,setspace,multirow}
\usepackage{graphicx}
\usepackage[top=2.4cm,left=2.4cm,top=2.4cm,bottom=2.4cm,includefoot]{geometry}
\usepackage{natbib}
\usepackage{caption}
\usepackage{pdflscape}

\bibliographystyle{bes}

@@

\textbf{Authors:} Dominique Gravel$^{1,2,*}$, Ben Baiser$^{3}$, Jennifer A. Dunne$^{4}$, Jens-Peter Kopelke$^{5}$, Neo
Martinez$^{6}$, Tommy Nyman$^{7}$, Timoth\'ee Poisot$^{2,8}$,  Spencer Wood$^{9}$, Daniel B. Stouffer$^{10}$, Jason Tylianakis$^{10,11}$ Tomas Roslin$^{12}$,\\

@@

3: \\

@@

5: Senckenberg Research Institute and Natural History Museum, Senckenberganlage 25, D-60325 Frankfurt am Main, Germany\\

@@

10: University of Canterbury at Christchurch, School of Biological Sciences\\

11:\\

12: Department of Ecology, Swedish University of Agricultural Sciences, Box 7044, 750 07 Uppsala, Sweden\\

@@

Biogeography has historically focused on the spatial distribution and
abundance of species, neglecting variation in the way species interact with
each other. Models of species distribution and of interactions have
historically taken different paths. Here, we plea for an integrated approach,
adopting the view that community structure is best represented as an
ecological network of interactions. We outline conceptual approach, suggesting
that the ecological niche may be redefined to encompass the effect of the
environment on species distribution (the Grinnellian dimension of the niche)
and on the ecological interactions among them (the Eltonian dimension).
Starting from this novel concept, we develop a quantitative theory to explain
turnover of interactions in space and time – i.e. a novel approach to
interaction distribution modelling. We decompose the probability with which
two species interact with each other in two parts: the probability with which
the species co-occur at a given location, and the probability that if they do
so, they will interact. Both of these probabilities could be conditional on
the environment. We apply this novel framework to a large data set of
host–parasite interactions across Europe and find that two aspects of the
environment (temperature and precipitation) leave a strong imprint on species
co-occurrence, but not on the probability of local interactions. Even where
species co-occur, interaction proves a stochastic rather than deterministic
process, adding to variation in realized network structure. We also find that
a large majority of species pairs are never found together, thus precluding
any inferences regarding their probability to interact. Our framework provides
a first conceptual framework to explain the variation of network structure at
large spatial scales and opens new perspectives at the frontier between
species distribution modelling and community ecology.

@@

Community ecology is defined in most textbooks as \textit{the study of the
interactions that determine the distribution and abundance of organisms}
\citep{Krebs2001}. Despite a general consensus on this definition
(Scheiner2007), it is surprising that research on variation in community
structure has focused mostly on the turnover of species composition
\citep{Anderson2011}, neglecting variation in the way species interact with
each other \citep{Poisot2015a}. Given this omission, it is perhaps not
surprising that biogeographers are still struggling to establish whether
interactions actually impact the distribution of species \citep{Wisz2012,
Kissling2012}. Recent attempts at accounting for interactions in species
distribution models \citep{Pollock2014, Pelissier2013} have brought some
methodological advances. Yet, these techniques are still based on a ’species-
based’ approach to communities, where interactions are merely treated as fixed
covariates affecting distribution.

As a more explicit description of interaction structure among species, the
network approach offers a convenient representation of communities. Species
are represented as nodes and interactions by links. To date, studies of
network diversity have mostly been concerned with the distribution of
interactions within locations, and less so with the variation among locations
\citep{Dunne2005, Bascompte2007, Ings2007, Kefi2012}. However, there is now
ample evidence that interaction networks vary in space and time
\citep{Poisot2012, Albouy2014, Poisot2016, Trojelsgaard2015}. Metacommunity
theory provides explanation for the variation in the distribution of the
different nodes \citep{Gravel2011c, Pillai2011}, but there is no explanation
to the joint variation of node and link occurrences. We urgently need a
conceptual framework to formalize these observations.

Given the historically different approaches to modelling the distributions of
species vs. interactions, there is an evident need to bring the two together.
Here, we plea for an integrated approach, adopting the view that community
structure is best represented as an ecological network of interactions. Based
on this idea, we propose a new description of the basic concept of the
ecological niche, now integrating the effect of the environment on species
distribution and on the ecological interactions among them. Starting from this
redefined concept, we develop a unified theory to explain turnover of
interactions in space and time. We first present the conceptual framework, and
then formalize it mathematically, using a probabilistic model to represent the
sampling of the regional pool of interactions. At the level of species pairs,
the statistical approach could be conceived as an interaction distribution
model. At the community level, the approach provides a likelihood-based method
to compare different hypotheses of network turnover. By applying this novel
framework to a large data set on host–parasite interactions across Europe, we
find that variation of the environment causes turnover of both species and
interactions. The network structure changes systematically across the
latitudinal gradient, with a peak of connectance at intermediate latitudes.

%=============================================================================%
\section*{The two dimensions of community structure}

The problem of community assembly is often formulated as \textit{how do we
sample a regional pool of species to constitute a local community?} This
question could be rewritten to address the problem of network assembly, as
\textit{how do we sample a regional pool of interactions to constitute a local
interaction network?} An illustration of this problem for a food web is
provided in Figure 1. The metaweb represents potential interactions among all
species that could be found in a given area. In this particular case, there
are 275 nodes, and 1173 links among plants (52 nodes), herbivores (96 nodes),
and parasitoids (127 nodes) from Northern Europe. An instance of a local
community is also illustrated, with 45 nodes and 93 interactions. Only 55.0\%
of all potential interactions are realized. Our objective here is to provide a
conceptual framework to explain the sampling of the regional pool of
interactions, along with a quantitative method to predict it. The problem
could be formalized sequentially by first understanding why only a fraction of
the species are co- occurring locally, and then why these species interact or
not.

There are multiple causes of spatial turnover in community composition. The
first and most-studied driver is the effect of variation in the abiotic
environment on species performance. Combined with specific responses in
demography, it generates variation among sites by selecting the locally
fittest species \citep{Leibold2004}. Stochasticity plays an additional role,
either through the inherently unpredictable nature of colonization and
extinction events (Hanski1999), or through strong non-linear feedbacks
generating alternative transients and equilibriums \citep{Chase2007,
Vellend2014}. Analyses of community turnover are usually performed with data
represented in a table with rows corresponding to sites (or measurements) and
columns to species. Metrics of beta diversity quantify the variance of this
community data \citep{Legendre2005}. Traditional approaches rely on measures
of dissimilarity among communities, such as the Jaccard or Bray–Curtis
indices. A more recent approach decomposes the total variation of the
community data into species and site contributions to beta diversity
\citep{Legendre2013}. Even though these methods compare whole lists of species
among sites or measurements, they remain fundamentally ’species-based’, since
they report the variation within columns. None of them explicitly considers
the variation of associations (i.e., of pairs or higher- order motifs –
\citealt{Stouffer2007}).

The “niche” is by far the dominant concept to explain species distributions
and community assembly, from the local to the global scale. Following
\citealt{Hutchinson1957}, the niche is viewed as the set of environmental
conditions allowing a population to establish and persist (see also
\citealt{Holt2009}. Community turnover arises as a result of successive
replacement of species along an environmental gradient, in agreement with the
Gleasonian view of communities \citep{Gleason1926}. The concept is
straightforward to operationalize with species distribution models, as it maps
naturally on available distributional and environmental data. In consequence,
a vast array of statistical tools have been developed to implement it (e.g.
Biomod \citealt{Thuiller2003}, MaxEnt \citealt{Phillips2006}). It is however
much harder to account for ecological interactions with this approach
(Peterson2011). As such, these interactions are often viewed as externalities
constraining or expanding the range of environmental conditions required for a
species to maintain a viable population \citep{Pulliam2000, Soberon2007}.

As mentioned above, in the network approach to community structure, species
and interactions are represented by nodes and links, respectively.
Associations can also be represented by matrices in which entries represent
the occurrence or intensity of interactions among species (rows and columns).
Network complexity is then computed as the number of interactions (in the case
of binary networks) or interaction diversity (in the case of quantitative
networks, \citealt{Bersier2002}). Variability in community structure
consequently arises from the turnover of species composition, along with
turnover of interactions among pairs of species. The occurrence and intensity
of interactions could vary because of the environment, species abundance, and
higher-order interactions \citep{Poisot2015a}. Variation in community
composition was found independent of the variation of ecological interactions,
suggesting that these two components of network variability respond to
different drivers \citep{Poisot2012}.

Interestingly, the ecological network literature also has its own ’niche
model’ to position a species in a community (Williams2000). The niche of a
species in this context represents the multidimensional space of all of its
interactions. Each species is characterized by a niche position, an optimum
and a range over 3 to 5 different niche axes \citep{Williams2000, Eklof2013}.
The niche model of food web structure has successfully explained the
complexity of a variety of networks, from food webs to plant–pollinator
systems \citep{Allesina2008, Williams2010, Eklof2013}. This
conceptual framework is, however, limited to local communities, and does not
provide any explanation for the turnover of network structure along
environmental gradients.

@@

A more integrative description of the niche is key to understand spatial and
temporal turnover in community structure. Despite several attempts to update
the concept of the ecological niche, ecologists have not moved far past the
“n-dimensional hypervolume” defined by Hutchinson. Despite its intuitive
interpretation and easy translation into species distribution models
\citep{Boulangeat2012, Blonder2014}, the concept has been constantly
criticized \citep{Hardin1960, Peters1991, Silvertown2004}, and several
attempts have been made to expand and reinforce it \citep{Pulliam2000,
Chase2003, Soberon2007, Holt2009, McInerny2012b}.

Part of the problem surrounding the niche concept has been clarified with the
distinction between Eltonian and Grinnellian definitions (Chase2003). The
Grinnellian dimension of the niche is the set of environmental conditions
required for a species to maintain a population in a location. The Grinnellian
niche is intuitive to apply, and constitutes the conceptual backbone of
species distribution models. The Eltonian niche, on the other hand, is the
effect of a species on its environment. This aspect of the niche is well known
by community ecologists, but is trickier to turn into predictive models.
Nonetheless, the development of the niche model of food web structure
\citep{Williams2000} and its parameterization \citep{Williams2010; Gravel2013}
made it more operational.

These perspectives are rather orthogonal to each other, which has resulted in
considerable confusion in the literature \citep{McInerny2012a}.
\citealt{Chase2003} attempted to reconcile them in their definition of the
niche: “[The niche is] the joint description of the environmental conditions
that allow a species to satisfy its minimum requirements so that the birth
rate of a local population is equal to or greater than its death rate along
with the set of per capita effects of that species on these environmental
conditions”. Their representation merges zero-net growth isoclines delimiting
the Grinnellian niche (when do populations persist) with impact vectors
delimiting the Eltonian niche (what is the per-capita impact). While this
representation has been very influential in local-scale community ecology (the
resource-ratio theory of coexistence, \citealt{Tilman1982}, it remains
impractical at larger spatial scales because of the difficulties to measure
it. The absence of any mathematical representation of the niche that could
easily be fitted to ecological data may explain why biogeographers are still
struggling to develop species distribution models that also consider
ecological interactions.

We propose to integrate the two perspectives of the niche with a visual
representation of both components. The underlying rationale is that, in
addition to the environmental constraints on demographic performance, any
organism requires resources to sustain its metabolic demands and reproduction.
Abiotic environmental axes are any non-consumable factors affecting the
demographic performance of an organism. Alternatively, the resource axes are
traits of the resources allowing interactions with the consumer. The niche
should therefore be viewed as the set of abiotic environmental factors (the
Grinnellian component) along with the set of traits (the Eltonian component)
allowing a population to establish and to persist at a location. Accordingly,
each species can be characterized by an optimal position in both the
environmental (x-axis) and the trait (y-axis) plane. The integrated niche is
then the hypervolume where interactions can occur and sustain a population.
This approach radically changes the representation of the niche, putting the
distributions and ecological interactions of species in the same formalism.
The limits of the niche axes could be independent of each other (as in the
example of Fig. 2), but they could also interact. For instance, the optimal
prey size for predatory fishes could decrease with increasing temperature
\citep{Lelong2015}, which would make diet boundaries functions of the
environment. The other way around, we could also consider that the growth rate
of the predator changes with the size of its prey items, thereby altering the
environmental boundaries.

%========================================================%
\section*{A probabilistic representation of ecological interactions networks in space}

We now formalize the integrated niche with a probabilistic approach to
interactions and distributions. We seek to represent the probability that an
interaction between species $i$ and $j$ occurs at location $y$. We define $L_{ijy}$ as a
stochastic variable, and are looking at the probability that this event
occurs, $P(L_{ijy})$. The occurrence of an interaction is dependent on the co-
occurrence of species $i$ and $j$. This argument might seem trivial at first, but
the explicit consideration of this condition in the probabilistic
representation of ecological interactions will prove fundamental to understand
their variation. We define $X_{iy}$ as a stochastic variable representing the
occurrence of species $i$ at location $y$, and similarly $X_{ijy}$ the co-occurrence
of species $i$ and $j$. The quantity we seek to understand is the probability of a
joint event, conditional on the set of environmental conditions $E_y$:

@@

Or simply said, the probability of observing both species $i$ and $j$, and an
interaction between $i$ and $j$. This probability could be decomposed in two parts
using the product rule of probabilities:

@@

The left term is the probability of observing the two species co-occurring at
location $y$. It corresponds to the Grinnellian dimension of the niche. The
right term is a conditional probability, representing the probability that an
interaction occurs between species $i$ and $j$, given their set of traits $T$ and
that they are co-occurring. Above, we referred to this entity as the metaweb,
corresponding to the Eltonian dimension of the niche . Below, we will see how
this formalism can be directly fitted to empirical data. But before turning to
an application, we will discuss the interpretation of different variants of
these two terms.

@@

There are several variants to the co-occurrence probability, representing
different hypotheses about temporal and spatial variation in network structure
(see the explicit formulations in Table 1). The simplest model relates the
probability of co-occurrence directly to the environment, $P(X_{ijy} |E_y)$. In
this situation there are no underlying assumptions about the ecological
processes responsible for co-occurrence. It could arise because of an impact
of ecological interactions on distributions \citep{Pollock2014} or, alternatively,
because of environmental requirements shared between i and j. In the former
case, species are not independent of each other and the conditional occurrence
must be accounted for explicitly, $P(X_{ijy} |E_y)=P(X_{iy} |E_y,X_{jy})P(X_{jy} |E_y)$.
In the latter case, species are independent and only the marginal occurrence
must be accounted for, $P(X_{ijy} |E_y)=P(X_{iy} |E_y)P(X_{jy} |E_y)$.

The co-occurrence probability itself could depend on ecological interactions.
This should be viewed as the realized component of the niche (i.e. the
distribution when accounting for species interactions). Direct pairwise
interactions such as competition, facilitation, and predation have long been
studied for their impact on co-distribution (e.g. \citealt{Diamond1976, Connor1980,
Gotelli2000}. Second- and higher-order interactions (e.g. trophic cascades)
could also affect co-occurrence. Co-occurrence of multiple species embedded in
ecological networks is however a topic of its own, influenced by the network
topology and species richness \citep{Cazelles2015}. Not only direct interactions do
influence co-occurrence, but indirect interactions as well (e.g. plant species
sharing an herbivore could repulse each other in space. The impact of direct
interactions and first-order indirect interactions on co-occurrence tends to
vanish with increasing species richness in the community. Further, co-
occurrence is also influenced by the covariance of interacting species to an
environmental gradient \citep{Cazelles2016}. Because of the complexity of relating
co-occurrence to the structure of interaction networks, we will here focus on
the variation of interactions and not on their distribution, and leave this
specific issue for the Discussion and future research.

@@

There are also variants of the metaweb. First, most documented metawebs have
thus far considered ecological interactions to be deterministic, not
probabilistic (e.g. \citealt{Havens1992, Woods2015}). Species are assumed to interact
whenever they are found together in a location, independently of their
abundance and the environment. In other words, $P(L_{ijy}|X_{ijy}=1)$ equals 1
andit equals 0 if $P(L_{ijy}|X_{ijy}=0)$. This approach might be a reasonable
approximation if the scale of sampling and inference is so large that the
probability of observing at least one interaction converges to unity and the
only variation in the networks considered arises from species distributions.

Ecological interactions could also vary with the environment, so that $P(L_{ijy}
|E_y)$. Although it is not common to see a conditional representation of
pairwise ecological interactions, experimental studies have revealed them to
frequently be sensitive to the environment. For instance, \citep{Mckinnon2010} shown
that predation risks of shorebirds vary at the continental scale, decreasing
from the south to the north. It is also common to see increasing top-down
control with temperature (e.g. \citealt{Shurin2012, Gray2016}). Effects of the
environment on interactions propagate up the community and influence network
structure \citep{Woodward2010; Petchey2010}.

We now turn to an illustration of the framework with the analysis of an
empirical dataset of host–parasite networks sampled throughout south–north
gradient in continental Europe. The focal system consists of local food webs
of willows (genus Salix), their galling insects, and the natural enemies of
these gallers. Targeting this system, we ask: i) how much does network
structure vary across the gradient, and ii) what is the primary driver of
network turnover across the gradient?

Communities of willows, gallers, and galler enemies are species-rich and
widely distributed, with pronounced variation in community composition across
space. The genus Salix includes over 400 species, most of which are shrubs or
small trees \citep{Argus1997}. The genus is common in most habitats across the
Northern Hemisphere \citep{Skvortsov1999}. Willows support a highly diverse
community of herbivorous insects, with one of the main herbivore groups being
gall-inducing sawflies (Hymenoptera: Tenthredinidae: Nematinae: Euurina
\citep{Kopelke1999}. Gall formation is induced by sawfly females during
oviposition, and includes marked manipulation of host-plant chemistry by the
galler \citep{Nyman2000}. The enemy community of the gallers includes nearly
100 species belonging to 17 insect families of four orders
\citep{Kopelke2000}. These enemies encompass two main types: inquiline larvae
(Coleoptera, Lepidoptera, Diptera, and Hymenoptera) feed primarily on gall
tissue, but typically kill the galler larva in the process, while parasitoid
larvae (representing many families in Hymenoptera) kill the galler larvae by
direct feeding \citep{Kopelke2003}. In terms of associations between the
trophic levels, phylogeny-based comparative studies have demonstrated that
galls represent “extended phenotypes” of the gallers, meaning that gall form,
location, and chemistry is determined mainly by the galling insects and not by
their host plants \citep{Nyman2000}. Because galler parasitoids have to
penetrate a protective wall of modified plant tissue in order to gain access
to their victims, gall morphology has been inferred to strongly affect the
associations between parasitoids and hosts \citep{Nyman2007}. Thus, the set of
parasitoids attacking each host is presumptively constrained by the form,
size, and thickness of its gall.

Local realizations of the willow–galler–enemy network were reconstructed from
community samples collected between 1982 and 2010. During this period, willow
galls were collected at 370 sites across Central and Northern Europe. Sampling
was conducted in the summer months of June and/or July, i.e., during the later
stages of larval development. Galler species were identified on the basis of
willow host species and gall morphology, as these are distinct for each sawfly
species. At each site, galls were randomly collected from numerous willow
individuals in an area of about 0.1–0.3 km2. Most sites were visited only
once, with a total of 641 site visits across the 370 sites. GPS coordinates
were recorded for each location; for our present analyses, the annual mean
temperature and precipitation were obtained from WorldClim using the R package
raster \citep{Hijmans2015}. While other covariates could have been considered,
we assumed mean temperature and annual precipitation to be representative of
the most important axes of the European climate, and more easily interpretable
than reduced variables obtained by e.g. PCA.

The methods used for rearing natural enemies from the galls have been
previously described by \citep{Kopelke1999, Kopelke2003}. In brief, galls were
opened to score the presence of galler or parasitoid/inquiline larvae. Enemy
larvae were classified to preliminary morphospecies, and the identity of each
morphospecies was determined by connecting them to adults emerging after
hibernation. The galls were reared by storing single galls in small glass
tubes \citep{Kopelke1985a}. Hibernation of galls containing enemies took place
either within the glass tubes or between blotting paper in flowerpots filled
with clay granulate or a mixture of peat dust and sand. These pots were stored
over the winter in a roof garden and/or in a climatic chamber. In most cases,
the matching of larval morphospecies with adult individuals emerging from the
rearings allowed the identification of the natural enemies to the species
level. Nonetheless, in some cases, individuals could only be identified to one
of the (super)families Braconidae, Ichneumonidae, and Chalcidoidea. This was
particularly the case when only remains of faeces, vacant cocoons of
parasitoids, and/or dead host larvae were found, as was the case when
parasitoids had already emerged from the gall. As a result, the largest taxon
in the data set, “Chalcidoidea indeterminate,” represents a superfamily of
very small parasitoids that are hard to distinguish.

In total, 146,622 galls from 52 Salix taxa were collected for dissection and
rearing. These galls represented 96 galler species, and yielded 42,133
individually-identified parasitoids and inquilines. Of these, 25,170 (60\%)
could be identified to the species level. Overall, 127 parasitoid and
inquiline taxa were distinguished in the material. Data on host associations
within subsets of this material have been previously reported by Kopelke1999
and by Nyman2007, whereas the current study represents the first analysis of
the full data set from a spatial perspective.

Computing the probability of observing an interaction involves fitting a set
of binomial models and collecting their estimated probabilities. For the sake
of illustration, we considered second-order generalized linear models –
whereas more sophisticated fitting algorithms (e.g. GAM or Random Forest)
could equally well be used, as long as the algorithm can estimate the
probability for each observation. The data consist of a simple (albeit large
and full of zeros) table with the observation of each species, $X_{iy}$ and $X_{jy}$,
their co-occurrence, $X_{ijy}$, the observation of an interaction $L_{ijy}$, and
environmental co-variates $E_y$. Thus, there is one row per pair of species per
site. We considered that an absence of a record of an interaction between co-
occurring species at a site means a true absence (see below for a discussion
on this issue).
